---
title: Interpretation of effects
---

```{r setup}
library(effectclass)
library(INBOtheme)
library(knitr)
library(tidyverse)
```

We compare the 95% credible intervals with a reference, upper and lower threshold to classify the strength of the effect into 10 classes.
The change of a linear trend is converted into a change over the length of the data.
The change of an index is the actual change between the two years.
The reference is set to 0 (no change).
The credible interval of a significant effect does not contain 0.
We selected a change of -25% (3/4 of the initial value) as the lower threshold.
We use the complement^[$\log(3/4) = -0.2877$ and $\log(4/3) = 0.2877$] of that (+33% or 4/3 of the initial value) as the upper threshold.
A -25% or +33% change over 5 years is equivalent to an average yearly change of -5.6% or +5.9% in case of a linear trend.

```{r tbl-threshold-yearly}
#| tbl-cap: Average yearly change over a period required to yield a total
#|   decrease to 3/4 or total increase to 4/3 of the initial value.
data.frame(
  duration = c(1, 5, 10, 12, 20, 30)
) |>
  mutate(
    decrease = log(0.75) / duration, increase = -log(0.75) / duration,
    across(c("decrease", "increase"), ~sprintf("%+.2f%%", 100 * (exp(.x) - 1)))
  ) |>
  kable()
```

Below are the symbols, interpretations and rules for each of the 10 classes.
Fig. \@ref(fig:effect-class) illustrates these rules.

- `++` **strong increase**: A significant positive trend and significantly stronger than the upper threshold.
- `+~` **moderate increase**: A significant positive trend and significantly weaker than the upper threshold.
- `+` **increase**: A significant positive trend, not significantly different from the upper threshold.
- `~` **stable**: A non-significant trend and significantly between the lower and upper threshold.
- `-` **increase**: A significant negative trend, not significantly different from the lower threshold.
- `-~` **moderate decrease**: A significant negative trend and significantly weaker than the lower threshold.
- `--` **strong decrease**: A significant negative trend and significantly stronger than the lower threshold.
- `?+` **potential increase**: A non-significant trend, significantly above the lower threshold.
- `?-` **potential decrease**: A non-significant trend, significantly above the upper threshold.
- `?` **unknown**: A non-significant trend, both the lower and upper threshold are probable.

One of the benefits is that we distinguish `~` (stable) and `?` (unknown). 
Both are not significant.
The main difference between both cases is the uncertainty.
We set the thresholds at important changes.
If the uncertainty is large, then the credible interval contains both the lower and the upper threshold.
So we have no clue what is happening, hence the unknown class.
If the uncertainty is small, then the credible interval contains neither the lower nor the upper threshold.
In this case we do known that the change is less extreme that the thresholds.
So if there is a change, it will be smaller than important changes (the thresholds).
This is informative, even though the change is not significant.

```{r fig-effect-class}
#| fig.cap: Illustration of the classification of effect based on their credible
#|   interval.
#|   The dashed line is the reference (no effect). The dotted lines indicate the
#|   lower and upper threshold.
threshold <- log(4 / 3)
data.frame(
  x = c(0, 1 / 2, -1 / 2, 1.5, 1, 0.5, -1.5, -1, -0.5, 0) * threshold,
  sd = c(rep(1.4, 3), rep(0.49, 7)) * threshold / qnorm(0.95)
) |>
  mutate(
    lcl = qnorm(0.05, mean = .data$x, sd = .data$sd),
    ucl = qnorm(0.95, mean = .data$x, sd = .data$sd),
    class = classification(.data$lcl, .data$ucl, threshold = threshold)
  ) |>
  ggplot(aes(x = class, y = x, ymin = lcl, ymax = ucl)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_hline(yintercept = c(-1, 1) * threshold, linetype = 3) +
  stat_fan(aes(link_sd = sd), geom = "rect") +
  scale_y_continuous(labels = change_labels, breaks = change_breaks(n = 3, extra = threshold)) +
  coord_flip() +
  theme(axis.title = element_blank())
```

## Overview tables

Part \@ref(p:summarising-tables) contains several tables with effects listing the results for all species with sufficient data.
Chapter \@ref(ch:yearly-change) displays the average yearly change in seasonal population average assuming a linear trend.
Section \@ref(s:yearly-change-12) contains the results using only the last twelve years of the data, while section \@ref(s:yearly-change-tot) uses the full data of the species.
The tables contain the scientific name of the species, the class of the effect (see section \@ref(s:interpretation)) and the estimated change and its 95% credible interval.
The tables are sorted along a decreasing trend.
Species with uncertain trend are listed at the bottom of the tables.
The number is chapter \@ref(ch:total-change) refer to exactly the same trends as in chapter \@ref(ch:yearly-change), expect that the trends are expressed as a total change over the range of the data instead of a an average yearly change.
Please note that when using the full dataset, the time range depends on the species.

Chapter \@ref(ch:wintermaxima) displays results using the winter maxima.
Section \@ref(s:wintermaxima) contains the average winter maximum over the last five year.
These tables contain both a detailed number and a number for publication.
The detailed number is the actual estimate from the model.
The number for publication is a rounded version of the detailed number, with the level of rounding depending on the uncertainty of the detailed number.
We strongly recommend to use the number of publication including the credible interval.
Section \@ref(s:wintermaxima-delta) estimate the relative change in average winter maxima over the last two five-year periods.
This changed is classified along the guidelines in section \@ref(s:interpretation).

## Figures

Each modelled species gets its own chapter with results in part \@ref(p:species).
The results are split into sections per group of sites.
All results are display in a standardized graphical format without comments or interpretation.

### Seasonal average

The results start with a figure showing the estimated seasonal average number of birds.
This is the estimated monthly population, corrected for the overall seasonal pattern.
The line displays the point estimate for each year.
This is the most likely value for the total number of birds in an average month.
The three ribbons display the uncertainty around this point estimate.
They are, from small/dark to wide/light, the 30%, 60% and 90% credible intervals.

### Indices and index raster

An index is a change compared to a baseline.
This baseline is typically the estimate for some reference winter.
E.g we use 2014 as a baseline and compare 2015 with 2013 or 2016 with 2014.
However we cannot use the figure with 2014 as baseline to compare 2015 with 2016.
For that we need a figure with either 2015 or 2016 as baseline.

The last two plots of the section display the index using the first and last available year as the reference.
Plotting all other years would require a large amount of figures.
Therefore we summarise the indices for all possible combinations on a raster.
The x axis holds the winter we want to interpret.
The y axis holds the reference winter.
The dots given the relative change from the baseline (y axis) to the other winter (x axis).
Their colour indicates the strength of the change.
Stronger changes have darker dots, white dots indicate no change.
Red dots indicate a decrease from the baseline, blue dots an increase.
A baseline with all red (blue) dots indicates the winter with the largest (smallest) numbers.
The shape of the dots indicates the classification of the effect (see section \@ref(s:interpretation)).
The equivalent positive and negative classes are collapsed into a single class. 
`++` and `--` become `**`, `+` and `-` become `*`, `+~` and `-~` become `*~` and finally `?+` and `?-` become `?*`.
Informative dots (significant or non-significant but stable) get solid shapes.
Uncertain estimates get an open shape.
