```{r mt-{{id}}-setup}
if (interactive()) {
  this_id <- sample(mt$id, 1)
} else {
  this_id <- "{{id}}"
}
mt %>%
  filter(id == this_id) -> selection
index %>%
  filter(id == this_id) -> this_index
wintermax %>%
  filter(id == this_id) -> this_wm
avg_wintermax %>%
  filter(id == this_id) %>%
  mutate(end = max(this_wm$Parameter), start = end - 4) -> this_avg_wm
if (max(selection$data[[1]]$UCL, na.rm = TRUE) > 1e4) {
  selection$data[[1]] %>%
    mutate_at(c("Estimate", "LCL", "UCL"), "/", 1e3) -> selection$data[[1]]
  this_index %>%
    mutate_at(vars(starts_with("p")), "/", 1e3) -> this_index
  this_wm %>%
    mutate_at(c("Estimate", "LCL", "UCL"), "/", 1e3) -> this_wm
  this_avg_wm %>%
    mutate_at(vars(starts_with("p")), "/", 1e3) -> this_avg_wm
  extra_lab <- "(x 1000)"
} else {
  extra_lab <- ""
}
location_effect %>%
  filter(id == this_id) -> this_location
imputations  %>%
  filter(id == this_id) -> this_imputations
```

`r selection$title`

### Monthly totals

(ref:mt-{{id}}) Monthly totals as subplots for _`r selection$scientific_name[1]`_ in `r selection$LocationGroup[1]`.

```{r mt-{{id}}-plot, fig.cap = "(ref:mt-{{id}})"}
selection$data[[1]] %>%
  mutate(type = str_remove(type, " total") %>%
           factor(levels = c("observed", "imputed"))) %>%
  ggplot(aes(x = winter, y = Estimate, ymin = LCL, ymax = UCL)) +
  geom_ribbon(alpha = 0.2, aes(group = type)) +
  geom_line(aes(colour = type)) +
  facet_wrap(~month, scales = "free_y") +
  scale_y_continuous(paste("Monthly total after imputation", extra_lab), 
                     limits = c(0, NA)) +
  theme(axis.title.x = element_blank())
```

(ref:mt-{{id}}-2) Imputed montly totals in a single plot for _`r selection$scientific_name[1]`_ in `r selection$LocationGroup[1]`.

```{r mt-{{id}}-plot2, fig.cap = "(ref:mt-{{id}}-2)"}
selection$data[[1]] %>%
  filter(type == "imputed total") %>%
  ggplot(aes(x = winter, y = Estimate, ymin = LCL, ymax = UCL)) +
  geom_ribbon(alpha = 0.2, aes(fill = month)) +
  geom_line(aes(colour = month)) +
  scale_y_continuous(paste("Monthly total after imputation", extra_lab), 
                     limits = c(0, NA)) +
  theme(axis.title.x = element_blank())
```

(ref:mt-{{id}}-3) Montly totals (blue) combined with the yearly average (green) for _`r selection$scientific_name[1]`_ in `r selection$LocationGroup[1]`.

```{r mt-{{id}}-plot3, fig.cap = "(ref:mt-{{id}}-3)"}
selection$data[[1]] %>%
  filter(type == "imputed total") %>%
  ggplot(aes(x = winter)) +
  stat_fan(data = this_index, aes(y = Estimate, link_sd = SE), 
              fill = inbo.steun.geelgroen, link = "log") +
  geom_ribbon(aes(ymin = LCL, ymax = UCL), alpha = 0.2) +
  geom_line(aes(y = Estimate)) +
  geom_line(data = this_index, aes(y = Estimate), colour = inbo.steun.geelgroen) +
  facet_wrap(~month, scales = "free_y") +
  scale_y_continuous(paste("Monthly total after imputation", extra_lab), 
                     limits = c(0, NA)) +
  theme(axis.title.x = element_blank())
```

### Winter maxima

(ref:wm-{{id}}) Winter maximum after imputation for _`r selection$scientific_name[1]`_ in `r selection$LocationGroup[1]`.

```{r wm-{{id}}-plot, fig.cap = "(ref:wm-{{id}})"}
ggplot(this_wm, aes(x = Parameter, y = Estimate, ymin = LCL, ymax = UCL)) +
  geom_ribbon(alpha = 0.2) +
  geom_line() +
  scale_y_continuous(paste("Winter maximum after imputation", extra_lab)) +
  theme(axis.title.x = element_blank())
```

(ref:wm-{{id}}-2) Winter maximum after imputation with indication of the average over the last five years for _`r selection$scientific_name[1]`_ in `r selection$LocationGroup[1]`.

```{r wm-{{id}}-plot2, fig.cap = "(ref:wm-{{id}}-2)"}
tibble(
  Parameter = this_avg_wm$start:this_avg_wm$end,
  Estimate = this_avg_wm$Estimate,
  SE = this_avg_wm$SE
) -> this_avg_wm
ggplot(this_wm, aes(x = Parameter, y = Estimate)) +
  stat_fan(data = this_avg_wm, aes(link_sd = SE), fill = inbo.steun.geelgroen) +
  geom_ribbon(aes(ymin = LCL, ymax = UCL), alpha = 0.2) +
  geom_line() +
  geom_line(data = this_avg_wm, colour = inbo.steun.geelgroen) +
  scale_y_continuous(paste("Winter maximum after imputation", extra_lab)) +
  theme(axis.title.x = element_blank())
```

```{r results = "asis"}
if (nrow(this_location)) {
  cat("### Importance of locations")
}
```

(ref:location-{{id}}) Histogram of the relative importance of the sites for _`r selection$scientific_name[1]`_ in `r selection$LocationGroup[1]`.

```{r location-{{id}}, fig.cap = "(ref:location-{{id}})"}
if (nrow(this_location)) {
  cat("### Importance of locations\n\n")
  this_location %>%
    ggplot(aes(x = Estimate)) + geom_histogram(binwidth = 0.05) +
    scale_x_continuous("Relative importance of a site", labels = percent) +
    ylab("number of sites")
}
```

```{r top-location-{{id}}, results = "asis"}
if (nrow(this_location)) {
this_location %>%
  arrange(desc(Estimate)) %>%
  head(20) %>%
  transmute(
    location, 
    effect = sprintf("%.1f%% (%.1f%%; %.1f%%)", 100 * Estimate, 100 * LCL, 
                     100 * UCL)) %>%
  pandoc.table(
    caption = sprintf("Most important sites for _%s_ in %s.", 
                      selection$scientific_name[1],
                      selection$LocationGroup[1]), 
    split.cells = 60, split.table = Inf)
}
```

### Imputations

```{r imputations-{{id}}, warnings = FALSE, echo = FALSE, message = FALSE}
if (nrow(this_imputations)) {
  this_imputations %>%
    group_by(location) %>%
    nest() %>%
    mutate(plot = map2(
      data, location,
      ~ggplot(.x, aes(x = year, y = Estimate, ymin = LCL, ymax = UCL)) +
        geom_errorbar(data = filter(.x, !is.na(LCL))) +
        geom_line() +
        geom_point(aes(colour = factor(is.na(LCL), levels = c(TRUE, FALSE), 
                                       labels = c("observed", "imputed")))) +
        facet_wrap(~month, scales = "free_y") +
        labs(y = "count", title = .y, colour = "status") +
        theme(axis.title.x = element_blank())) %>%
        map(print)
    ) -> junk
}
```
