# (PART) Summarising tables {-}

```{r helper-functions}
estimate <- function(mean, round = 1) {
  round(exp(mean), round)
}
publication <- function(mean, magnitude) {
  magnitude <- magnitude - 1
  round(exp(mean) / 10 ^ magnitude) * 10 ^ magnitude
}
```

# Average yearly change based on monthly totals

## Yearly change over the last twelve years

```{r monthly-trend-12-yearly, results = "asis"}
results %>%
  filter(ModelType == "imputed trend: Total ~ Year + Month", Duration == 12, 
         Parameter == "cYear") %>%
  mutate(Class = classification(LCL, UCL, log(0.75) / Duration[1])) %>%
  mutate_at(c("Estimate", "LCL", "UCL"), change) %>%
  transmute(
    Area = LocationGroup, Species = scientific_name, 
    Estimate, Class,
    Change = sprintf("%+.0f%% (%+.0f%%; %+.0f%%)", 
                     ifelse(Estimate > 1e5, Inf, Estimate), LCL, 
                     ifelse(UCL > 1e5, Inf, UCL))
  ) %>%
  arrange(Area, Class, desc(Estimate)) %>%
  mutate(Class = format(Class, type = "markdown")) %>%
  select(-Estimate) %>%
  group_by(Area) %>%
  nest() -> tab
junk <- sapply(
  seq_along(tab$Area),
  function(i) {
    cat("###", as.character(tab$Area[i]), "\n\n")
    print(kable(
      tab$data[[i]], format = "pandoc", align = "lcc",
      label = paste0("monthly-trend-12-yearly", 
                     str_replace(tab$Area[i], " ", "-"), sep = "-"),
      caption = paste("Average yearly change over the last twelve year based on
                      monthly totals for", tab$Area[i])))
    cat("\n\n")
  }
)
```

## Yearly change over the entire dataset

```{r monthly-trend-all-yearly, results = "asis"}
results %>%
  filter(ModelType == "imputed trend: Total ~ Year + Month", 
         Parameter == "cYear") %>%
  group_by(SpeciesGroup) %>%
  filter(Duration == max(Duration)) %>%
  ungroup() %>%
  group_by(Duration) %>%
  mutate(Class = classification(LCL, UCL, log(0.75) / Duration[1])) %>%
  mutate_at(c("Estimate", "LCL", "UCL"), change) %>%
  transmute(
    Area = LocationGroup, Species = scientific_name, 
    Class, Estimate,
    Change = sprintf("%+.0f%% (%+.0f%%; %+.0f%%)", 
                     ifelse(Estimate > 1e5, Inf, Estimate), LCL, 
                     ifelse(UCL > 1e5, Inf, UCL))
  ) %>%
  arrange(Area, Class, desc(Estimate)) %>%
  mutate(Class = format(Class, type = "markdown")) %>%
  select(-Estimate) %>%
  group_by(Area) %>%
  nest() -> tab
junk <- sapply(
  seq_along(tab$Area),
  function(i) {
    cat("###", as.character(tab$Area[i]), "\n\n")
    print(kable(
      tab$data[[i]], format = "pandoc", align = "lcc",
      label = paste0("monthly-trend-all-yearly-total", 
                     str_replace(tab$Area[i], " ", "-"), sep = "-"),
      caption = paste("Average yearly change over the entire dataset based on
                      monthly totals for", tab$Area[i])))
    cat("\n\n")
  }
)
```

# Total linear change based on monthly totals

## Total change over the last twelve years

```{r monthly-trend-12-yearly-total, results = "asis"}
results %>%
  filter(ModelType == "imputed trend: Total ~ Year + Month", Duration == 12, 
         Parameter == "cYear") %>%
  mutate(
    Class = classification(LCL, UCL, log(0.75) / Duration[1]),
    Estimate = total_change(Estimate, Duration),
    LCL = total_change(LCL, Duration),
    UCL = total_change(UCL, Duration)
  ) %>%
  transmute(
    Area = LocationGroup, Species = scientific_name, 
    Class, Estimate,
    Change = sprintf("%+.0f%% (%+.0f%%; %+.0f%%)", 
                     ifelse(Estimate > 1e5, Inf, Estimate), LCL, 
                     ifelse(UCL > 1e5, Inf, UCL))
  ) %>%
  arrange(Area, Class, desc(Estimate)) %>%
  mutate(Class = format(Class, type = "markdown")) %>%
  select(-Estimate) %>%
  group_by(Area) %>%
  nest() -> tab
junk <- sapply(
  seq_along(tab$Area),
  function(i) {
    cat("###", as.character(tab$Area[i]), "\n\n")
    print(kable(
      tab$data[[i]], format = "pandoc", align = "lcc",
      label = paste0("monthly-trend-12-total", 
                     str_replace(tab$Area[i], " ", "-"), sep = "-"),
      caption = paste("Average total change over the last twelve year based on
                      monthly totals for", tab$Area[i])))
    cat("\n\n")
  }
)
```

## Total linear change over the entire dataset

```{r monthly-trend-all-yearly-total, results = "asis"}
results %>%
  filter(ModelType == "imputed trend: Total ~ Year + Month", 
         Parameter == "cYear") %>%
  group_by(SpeciesGroup) %>%
  filter(Duration == max(Duration)) %>%
  group_by(Duration) %>%
  mutate(
    Class = classification(LCL, UCL, log(0.75) / Duration[1]),
    Estimate = total_change(Estimate, Duration),
    LCL = total_change(LCL, Duration),
    UCL = total_change(UCL, Duration)
  ) %>%
  transmute(
    Area = LocationGroup, Species = scientific_name, 
    Class, Estimate,
    Change = sprintf("%+.0f%% (%+.0f%%; %+.0f%%)", 
                     ifelse(Estimate > 1e5, Inf, Estimate), LCL, 
                     ifelse(UCL > 1e5, Inf, UCL))
  ) %>%
  arrange(Area, Class, desc(Estimate)) %>%
  mutate(Class = format(Class, type = "markdown")) %>%
  select(-Estimate) %>%
  group_by(Area) %>%
  nest() -> tab
junk <- sapply(
  seq_along(tab$Area),
  function(i) {
    cat("###", as.character(tab$Area[i]), "\n\n")
    print(kable(
      tab$data[[i]], format = "pandoc", align = "lcc",
      label = paste0("monthly-trend-all-total", 
                     str_replace(tab$Area[i], " ", "-"), sep = "-"),
      caption = paste("Average total change over the entire dataset based on
                      monthly totals for", tab$Area[i])))
    cat("\n\n")
  }
)
```

# Winter maxima

## Average winter maximum over the last five years

```{r wintermax, results = "asis"}
results %>%
  filter(ModelType == "imputed average: Total ~ cPeriod", 
         Parameter == "(Intercept)") %>%
  mutate(
    UCL = ifelse(UCL > 20, Inf, UCL),
    Magnitude = floor(log10(exp(UCL))),
    Magnitude = Magnitude - (1.2 * 10 ^ Magnitude > exp(UCL)),
    pX = publication(Estimate, Magnitude),
    pL = publication(LCL, Magnitude),
    pU = publication(UCL, Magnitude)
  ) %>%
  mutate_at(c("Estimate", "LCL", "UCL"), estimate) %>%
  transmute(Area = LocationGroup, Species = scientific_name, Estimate,
            Detail = sprintf("%.0f (%.0f; %.0f)", Estimate, LCL, UCL),
            "For publication" = sprintf("%.0f (%.0f; %.0f)", pX, pL, pU)) %>%
  arrange(desc(Estimate)) %>%
  select(-Estimate) %>%
  group_by(Area) %>%
  nest() -> tab
junk <- sapply(
  seq_along(tab$Area),
  function(i) {
    cat("###", as.character(tab$Area[i]), "\n\n")
    print(kable(
      tab$data[[i]], format = "pandoc", align = "lcc",
      label = paste0("wintermax", 
                     str_replace(tab$Area[i], " ", "-"), sep = "-"),
      caption = paste("Average winter maximum over the last five years for", 
                      tab$Area[i])))
    cat("\n\n")
  }
)
```

## Change in average winter maximum over the last two five-year periods

```{r wintermax-change, results = "asis"}
results %>%
  filter(ModelType == "imputed average: Total ~ cPeriod", 
         Parameter == "cPeriod") %>%
  transmute(
    Area = LocationGroup, Species = scientific_name, 
    Class = classification(LCL, UCL, log(0.75)),
    Estimate,
    Change = sprintf("%+.0f%% (%+.0f%%; %+.0f%%)", 
                     ifelse(change(Estimate) > 1e5, Inf, change(Estimate)), 
                     change(LCL), ifelse(change(UCL) > 1e5, Inf, change(UCL)))
  ) %>%
  arrange(Area, Class, desc(Estimate)) %>%
  mutate(Class = format(Class, type = "markdown")) %>%
  select(-Estimate) %>%
  group_by(Area) %>%
  nest() -> tab
junk <- sapply(
  seq_along(tab$Area),
  function(i) {
    cat("###", as.character(tab$Area[i]), "\n\n")
    print(kable(
      tab$data[[i]], format = "pandoc", align = "lcc",
      label = paste0("wintermax-change", 
                     str_replace(tab$Area[i], " ", "-"), sep = "-"),
      caption = paste("Change in average wintermax over the last two five-year 
                      periods for", tab$Area[i])))
    cat("\n\n")
  }
)
```
