```{r {{id}}-set}
if (interactive()) {
  this_id <- sample(species_index$loc_id, 1)
} else {
  this_id <- "{{id}}"
}
start_year <- 0
end_year <- 0
```

```{r {{id}}-select, results = "asis"}
species_index %>%
  filter(loc_id == this_id) -> selection
cat(selection$title)
```

(ref:{{id}}-population) Yearly population average based on the monthly population total of _`r selection$scientific_name`_ in `r selection$LocationGroup`.

```{r {{id}}-population, fig.cap = "(ref:{{id}}-population)"}
if (length(selection$data)) {
selection$data[[1]] %>%
  filter(grepl("^total: ", Parameter)) %>%
  mutate(
    Parameter = str_remove(Parameter, "total: ") %>%
      as.integer()
  ) -> index
  if (nrow(index)) {
    if (max(qnorm(0.95, index$Estimate, index$SE)) > log(1e4)) {
      index %>%
        mutate(Estimate = Estimate - log(1e3)) -> index
      ylab <- "seasonal average (x 1000)"
    } else {
      ylab <- "seasonal average"
    }
    start_year <- min(index$Parameter, na.rm = TRUE)
    end_year <- max(index$Parameter, na.rm = TRUE)
    index %>%
      ggplot(aes(x = Parameter, y = exp(Estimate), link_sd = SE)) +
      stat_fan(link = "log") +
      geom_line() +
      scale_x_continuous(
        breaks = function(x) {
          if (diff(x) < 5) {
            x[1]:x[2]
          } else {
            pretty(x)
          }
        },
        expand = expand_scale(mult = c(0, 0.05), add = c(2, 0))
      ) +
      scale_y_continuous(ylab, limits = c(0, NA)) +
      theme(axis.title.x = element_blank())
  }
}
```

(ref:{{id}}-change) Pairwise changes among years of _`r selection$scientific_name`_ in `r selection$LocationGroup`.

```{r {{id}}-change, fig.cap = "(ref:{{id}}-change)"}
if (length(selection$data)) {
  selection$data[[1]] %>%
    filter(grepl("^index", Parameter)) %>%
    tidyr::extract(Parameter, into = c("To", "From"), convert = TRUE, 
                   regex = "index: ([[:digit:]]*)-([[:digit:]]*)") -> indices
  indices %>%
    mutate(Estimate = -Estimate) %>%
    rename(Old = From, From = To) %>%
    rename(To = Old) %>%
    bind_rows(indices) -> indices
  breaks <- c(100/99, 50/49, 20/19, 10/9, 5/4, 4/3, 3/2, 2, 3, 5, 10, 20)
  n <- sum(breaks <= exp(max(indices$Estimate))) + 1
  breaks <- tail(head(breaks, n), 4)
  if (length(breaks) == 1) {
    limits <- c(-1, 1) * log(breaks)
  } else {
    limits <- range(indices$Estimate)
  }
  breaks <- log(c(1 / rev(breaks), 1, breaks))
  ggplot(
    indices, 
    aes(y = From, x = To, colour = Estimate, ymin = LCL, ymax = UCL)
  ) +
    stat_effect(threshold = log(0.75)) +
    coord_fixed() +
    scale_x_continuous(breaks = pretty) +
    scale_y_continuous("Reference", breaks = pretty) +
    scale_colour_gradient2("Change", breaks = breaks, limits = limits,
                         labels = sprintf("%+.0f%%", change(breaks))) +
    scale_effect() +
    theme(
      axis.title.x = element_blank(),
      legend.key.size = unit(0.8, "lines")
    )
}
```

(ref:{{id}}-index) Yearly index based on the monthly population total of _`r selection$scientific_name`_ in `r selection$LocationGroup`. The index of `r start_year` is set to 100.

```{r {{id}}-index, fig.cap = "(ref:{{id}}-index)"}
indices %>%
  filter(From == min(From)) -> index
if (nrow(index)) {
  if (max(qnorm(0.95, index$Estimate, index$SE)) > log(1e2)) {
    index %>%
      mutate(Estimate = Estimate - log(1e3)) -> index
    ylab <- "Index (x 1000)"
    reference <- 1e-3
  } else {
    ylab <- "Index"
    reference <- 1
  }
  index %>%
    ggplot(aes(x = To, y = exp(Estimate), link_sd = SE)) +
    geom_hline(yintercept = reference, linetype = 2) +
    stat_fan(link = "log") +
    geom_line() +
    scale_x_continuous(
      breaks = pretty, 
      expand = expand_scale(mult = c(0, 0.05), add = c(2, 0))
    ) +
    scale_y_continuous(ylab, labels = percent) +
    theme(axis.title.x = element_blank())
}
```

(ref:{{id}}-index-last) Yearly index based on the monthly population total of _`r selection$scientific_name`_ in `r selection$LocationGroup`. The index of `r end_year` is set to 100.

```{r {{id}}-index-last, fig.cap = "(ref:{{id}}-index-last)"}
indices %>%
  filter(From == max(From)) -> index
if (nrow(index)) {
  if (max(qnorm(0.95, index$Estimate, index$SE)) > log(1e2)) {
    index %>%
      mutate(Estimate = Estimate - log(1e3)) -> index
    ylab <- "Index (x 1000)"
    reference <- 1e-3
  } else {
    ylab <- "Index"
    reference <- 1
  }
  index %>%
    ggplot(aes(x = To, y = exp(Estimate), link_sd = SE)) +
    geom_hline(yintercept = reference, linetype = 2) +
    stat_fan(link = "log") +
    geom_line() +
    scale_x_continuous(
      breaks = pretty, 
      expand = expand_scale(mult = c(0, 0.05), add = c(2, 0))
    ) +
    scale_y_continuous(ylab, labels = percent) +
    theme(axis.title.x = element_blank())
}
```

\clearpage
